---
title: "Introduction to the RivRetrieve R package"
author: 'Ryan M. Riggs and Simon Moulds'
date: 'August 17, 2023'
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{my-vignette}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding[utf8]{inputenc}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",warning=FALSE,message=FALSE,fig.width=7
)
```
## Here is a quick tutorial for RivRetrieve functions:

The `RivRetrieve` package is developed to efficiently access and download global river gauge data into an R environment. While numerous river data exists, our package is currently limited to daily measurements of river stage (meters) and discharge (cubic meters per second) for: Australia, Brazil, Canada, Chile, France, Japan, South Africa, the United Kingdom, and the United States.
```{r setup}
library(RivRetrieve)
# Amazon River near Obidos, Brazil
siteNumber='17050001'
discharge=brazil(site=siteNumber,variable='discharge')
stage=brazil(site=siteNumber,variable='stage')

plot(discharge$Date,discharge$Q, type='l',xlab='',ylab='Discharge (cms)')
```

RivRetrieve automatically outputs a dataframe containing a `Date` column of `datetime` values and either a `H` or `Q` `numeric` column for the stage and discharge variable, respectively. 

Stage is provided in meters and discharge is provided in cubic meters per second.


## Accessing the raw gauge data:
```{r}
raw=original(discharge)
```
```{r,echo=FALSE}
library(knitr)
kable(raw[1:5,],caption='Raw Brazilian gauge data')
```

## Site locations for all agencies can also be found:
```{r}
japanSites=japan(sites=TRUE)
```
```{r,echo=FALSE}
kable(japanSites[1:10,],caption='Example Japanese river gauge locations')
```

## Date defined retrieval
Specific time periods can also be accessed if the entire timeseries is not needed. This is particularly useful for japanese gauge data as it is time consuming to download.
```{r}
# Last annual timeseries
recent=uk(site="http://environment.data.gov.uk/hydrology/id/stations/3c5cba29-2321-4289-a1fd-c355e135f4cb",variable='stage',start_date = '2023-01-01',end_date = '2023-12-31')

plot(recent$Date,recent$H,type='l', xlab='',ylab='Stage (m)')
```
