---
title: "Introduction to the RivRetrieve R package"
author: 'Ryan M. Riggs and Simon Moulds'
date: "`r format(Sys.time(), '%B %d, %Y')`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{my-vignette}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding[utf8]{inputenc}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",warning=FALSE,message=FALSE,fig.width=7
)
```
## Here is a quick tutorial for RivRetrieve functions:

The `RivRetrieve` package is developed to efficiently access and download global river gauge data into an R environment. While numerous river data exists, our package is currently limited to daily measurements of river stage (meters) and discharge (cubic meters per second) for: Australia, Brazil, Canada, Chile, France, Japan, South Africa, the United Kingdom, and the United States.
```{r setup}
library(RivRetrieve)
# Amazon River near Obidos, Brazil
siteNumber='17050001'
discharge=brazil(site=siteNumber,variable='discharge')
stage=brazil(site=siteNumber,variable='stage')

plot(discharge$Date,discharge$Q, type='l',xlab='',ylab='Discharge (cms)')
```

RivRetrieve automatically outputs a dataframe containing a `Date` column of `datetime` values and either a `H` or `Q` `numeric` column for the stage and discharge variable, respectively. 

Stage is provided in meters and discharge is provided in cubic meters per second.


## Accessing the raw gauge data:
```{r}
raw=original(discharge)
```
```{r,echo=TRUE}
library(knitr)
kable(raw[1:5,],caption='Raw Brazilian gauge data')
```

## Site locations for all agencies can also be found:
```{r}
brazilSites=brazil(sites=TRUE)
```
```{r,echo=TRUE}
kable(brazilSites[1:10,],caption='Example Brazilian river gauge locations')
```

## Date defined retrieval
Specific time periods can also be accessed if the entire time series is not needed.
```{r}
# Last annual timeseries
siteNumber='17050001'
start='2023-01-01'
end='2023-01-31'
recent=brazil(site=siteNumber,variable='stage',start_date = start,end_date = end)

plot(recent$Date,recent$H,type='l', xlab='',ylab='Stage (m)')
```
